---
title: "Bayesian Regression Analyses"
author: "Christoph Völtzke"
date: "15 6 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results = "hide", message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(bain)
library(papaja)
library(tinylabels)
library(kableExtra)
library(RColorBrewer)
```

## Introduction to the data

In this file all analyses for different combinations of predictors is conducted. Not all of these analyses are conducted in the original manuscript file.
```{r, results = "hide", message=FALSE, warning=FALSE}
# Load the data - I mention in the text where I got the data from 
nba <- read.table("Data/nba.csv", sep=",", header=T)
```

```{r investiagte assumptions of data set,include=FALSE}
# not included as too much space in html
# if you want ot see the plots
# inspect the data and check the normality of the Residuals
hist(nba$Salary, main = "Distribution of Log transformed salary in 2022 season") # this variable is highly skewed, so we take the log of the salary to get a more normal distribution
hist(nba$PTS, main = "Distribution of Average Number of Points per game")
hist(nba$AST, main = "Distribution of Average Number of Assists per game")
hist(nba$DRB, main = "Distribution of Average Number of Defensive Rebounds per game")
hist(nba$STL, main = "Distribution of Average Number of Steals per game") # all other variables seem to follow a normal distribution so I do not have to expect severe violations of normality.
qqnorm(log(nba$Salary), main = 'Normal Q-Q plot Salary'); qqline(log(nba$Salary)) 
qqnorm(nba$PTS, main = 'Normal Q-Q plot Points'); qqline(nba$PTS)
qqnorm(nba$AST, main = 'Normal Q-Q plot Assists'); qqline(nba$AST)
qqnorm(nba$DRB, main = 'Normal Q-Q plot Defensive Rebounds'); qqline(nba$DRB)
qqnorm(nba$STL, main = 'Normal Q-Q plot Steals'); qqline(nba$STL) # next, I also checked for linearity.Also here, no severe violations can be observed. Therefore, we can continue with this data in the following analyses 
```

```{r}
hist(nba$Salary, main = "Distribution of Log transformed salary in 2022 season") # this variable is highly skewed, so we take the log of the salary to get a more normal distribution
hist(nba$PTS, main = "Distribution of Average Number of Points per game")
hist(nba$AST, main = "Distribution of Average Number of Assists per game")
hist(nba$DRB, main = "Distribution of Average Number of Defensive Rebounds per game")
hist(nba$STL, main = "Distribution of Average Number of Steals per game") # all other variables seem to follow a normal distribution so I do not have to expect severe violations of normality.
```

## 1 - specify prior distributions for your parameters
```{r set all initial prior and data sets}
set.seed(976) # set a seed for reproducability
# initial values are randomly chosen between 1 and 8 and are right now for 3 chains -  I also included matrices of initial values for a varying number of predictors
init2_1<- matrix(runif(4, min=1, max=8))
init2_2 <- matrix(runif(4, min=1, max=8))
init2_3 <- matrix(runif(4, min=1, max=8))
init2 <- rbind(matrix(init2_1,1,4),matrix(init2_2,1,4),matrix(init2_3,1,4))

init4_1 <- matrix(runif(6, min=1, max=8))
init4_2 <- matrix(runif(6, min=1, max=8))
init4_3 <- matrix(runif(6, min=1, max=8))
init4 <- rbind(matrix(init4_1,1,6),matrix(init4_2,1,6),matrix(init4_3,1,6))
                                                              
# uninformative priors for normal distribution - For now I set them to be uninformative as I don't have any knowledge about the size of the effects. In later stages I will try to incorporate historical data to make use of informative priors.
sigma02 <- c(1000,1000,1000); mu02 <- c(0,0,0)
sigma04 <- c(1000,1000,1000,1000,1000); mu04 <- c(0,0,0,0,0)

# uninformative priors for inverse gamma distribution, these to values are specified before and don't need to be entered in the function as I didn't intend to change them
alpha0 <- 0.001; beta0 <- 0.001

# specifying the data sets including the predictors of interest - As I am interested in the effect of different predictors on the salary I try out several combinations of predictors - All of these data frames are used in the functions so keep the right naming in mind
x2 <- cbind(nba$DRB,nba$AST)
x2.1 <- cbind(nba$PTS,nba$AST)
x2.2 <- cbind(nba$DRB,nba$STL)
x2.3 <- cbind(nba$PTS,nba$DRB)
x4 <- cbind(nba$PTS,nba$AST,nba$DRB,nba$STL)

y <- log(nba$Salary)
```

## 2./3.  Gibbs sampler & Metropolis-Hastings step

```{r}
# For full functions see Full_code.Rmd or the seperate functions
source("Functions/bayesian_regression.R")
```


## 4. - Assess convergence of the model

```{r}
# For full functions see Full_code.Rmd or the seperate functions
source("Functions/Convergence.R")
```

## 5. - Check a model assumption with PPP

```{r}
# For full functions see Full_code.Rmd or the seperate functions
source("Functions/Posterior_predictive_p_value.R")
```

## 6. - Obtain parameter estimates, credible intervals
```{r parameter estimates and credible intervals}
# the output includes the parameter estimates for each sampled parameter as well as the credible intervals and further descriptive statistics, which might be of interest
twopred <- bayesian_reg(y, x2, init2, sigma02, mu02, 10000, 40000, 3, 976)
twopred_O <- bayesian_reg(y, x2.1, init2, sigma02, mu02, 10000, 40000, 3, 976)
twopred_D <- bayesian_reg(y, x2.2, init2, sigma02, mu02, 10000, 40000, 3, 976)
twopred_OD <- bayesian_reg(y, x2.3, init2, sigma02, mu02, 10000, 40000, 3, 976)
fourpred <- bayesian_reg(y, x4, init4, sigma04, mu04, 10000, 40000, 3, 976)
```

## 7. - a.) Compare multiple model by means of DIC
```{r DIC}
# Calculating the DIC for all models of relevance based on the output of the bayesian_reg function
twopred$DIC
twopred_O$DIC
twopred_D$DIC
twopred_OD$DIC
fourpred$DIC
```

## 7- b.) Convergence

```{r}
convergencecheck(twopred$sampled_values_chains,50,976)
convergencecheck(fourpred$sampled_values_chains,50,976)
```

## 7. - c.) Bayes factor
```{r}
# Models to be tested
OLS4 <- lm(log(Salary) ~ PTS + AST + DRB + STL, data = nba)
OLS2 <- lm(log(Salary) ~ AST + DRB, data = nba)

set.seed(976)
# Using the bain package to obtain the bayes factor
# Bayes factor 1
bf1 <- bain(OLS4, hypothesis = "PTS > 0 & DRB > 0 & AST > 0 & STL > 0", standardize = TRUE); print(bf1)
# Bayes factor 2
bf2 <- bain(OLS4, hypothesis = "PTS > DRB > AST> STL", standardize = TRUE); print(bf2)
bf3 <- bain(OLS2, hypothesis = "AST> DRB", standardize = TRUE); print(bf3)
bf1
bf2
bf3
```

## 7. - d.) PPP

```{r}
ppp(fourpred$sampled_values,y,x4,3000,976)
ppp(twopred$sampled_values,y,x2,3000,976)
```

